---
title: "Analysis of mtcars dataset to infer the relationship mpg ~ transmission"
author: "Sathish Duraisamy"
date: "January 24, 2015"
output: html_document
---

# Context:
The Motor Trend, an automobiles magazien company would like the answers for
the following questions, looking at a particular dataset: mtcars.

    1. Is an automatic or manual transmission better for MPG?
    2. Quantify the MPG difference between automatic and manual transmissions

# Exploratory Data Analysis
We will take a look at:
    *the dimension of the dataset
    *how many variables are there and their names, classes (are there any factors)
    * first few observations of dataset
    *how are the vars correlated to each other (with a pairs plot)
       and also numerically with a cor() function call.
    *etc.,
```{r}
library(datasets)
data(mtcars)

dim(mtcars)
names(mtcars)
head(mtcars, 3)
```

# Coefficient Interpretation
```{r}
cor(mtcars$mpg, mtcars[, -c(1)])
```
See appendix for a plot of the variables: mpg vs am. The 'am' var DOES have a 
postive correlation (0.5998324) meaning: *as the tranmission type increases 
from 0(Auto) to 1(Manual) we see an increase in mpg*. We'll further
confirm and quantify this coef for the two categories ('Auto' and 'Manual') of am.

Convert 'am' to a factor var, and add it to dataset for ease with fitting & plotting
```{r}
mtcars$am <- as.factor(mtcars$am); levels(mtcars$am) <- c("Auto", "Manual")
```

# Question 1: Would Auto or Manual transmission give better MPG?
We will do a Student t-test with 95% confident interval to see which one category
is more likely.
```{r}
t.test(mtcars$mpg ~ mtcars$am, confid.level=0.975)
```
The p-value 0.001374 and estimated means of Auto vs Manual (24.39) suggest that
Manual transmission is more likely to give better MPG than Auto.

# Question 2: How to quantify the MPG diff between Auto and Manual transmissions?
To answer this, we will use 'Linear Regression' and fit multiple models:
first with both values of 'am', then with 'qsec' kept at a constant and seeing 
for 'am' vaues == 0 (Auto) and 'am' valus = 1 (Manual). We will then compare 
these models using 'anova'.
```{r}
fit1 <- lm(mpg ~ am, data=mtcars)
fit2 <- lm(mpg ~ I(factor(am)):qsec, data=mtcars)
summary(fit2)
```
# Executive Summary
From the summary(fit2) output, we could say, during the '1/4 mile time'(qsec), 
a car with Manual transmission gives 2.2911 MPG more while a car with Auto
transmission will only give 1.7815 MPG. 
See appendix for a Residual plot of this fitted model.

```{r}
anova(fit1, fit2)
```
Anova output supports our findings.

-------
#Appendix

## scatterplot of mtcars (mpg vs am) to visually explore how they are correlated
```{r}
data(mtcars)
plot(mtcars$am, mtcars$mpg, xlab="Transmission Type (0 = Auto, 1 = Manual)",
     ylab="Miles per gallon", main="mpg vs am:  correlation" )
```

## Residuals plotted
```{r}
residuals <- resid(fit1)
plot(mtcars$mpg, residuals, ylab="Residuals", 
     xlab="Miles Per Gallon", 
     main="mtcars mpg vs residuals") 
abline(0, 0)
```

--end--
